title: "Understanding entropy and mutual information applied to the face orientation tuning"
author: "Tancy"
date: "5 February 2015"
output: html_document
---
---

```{r loadLibary}
rm(list=ls())
library(plyr)
library(entropy)
library(ggplot2)
library(ggthemes)
library(reshape2)
library(gridExtra)
library(ks)
library(plot3D)
#options(device='quartz')

setwd("~/Dropbox/02_RikenProj/fMRI_Data_R/")
SigCh_OrientData <- read.csv('SigCh_OrientData.csv')

```


1st step: z normalized of the timeseies from individual voxel
```{r z_normalizeForSig}
RandomVox <- droplevels(subset(SigCh_OrientData, Subj=='S02'& orient!=7 & orient!=6 & ROIs!='rOTA' & ROIs!='lOTA'))
                              #& volume >= 1 & volume <=15
                              #& voxel %in% c(1,2,3, 7,8,25,45) )) # get 3 voxels info, remove blank baseline #,7,8,25,45
#head(RandomVox)
RandomVox_Zdata <- droplevels(ddply(RandomVox, .(voxel), transform,
                         zSigCh = scale(SignalChange)))
#hist(RandomVox_Zdata$zSigCh) # plot histogram

# writeCSVname
savefile <- "s02_MI_5orient.csv"
```

```{r z_normalizeForBeta}
# RandomVox <- droplevels(subset(Beta_OrientData, Subj=='S02'& ROIs=='rOFA' & orient!=7 & orient!=6 & volume >= 1 & volume <=20))
#                               #& voxel %in% c(1,2,3, 7,8,25,45) )) # get 3 voxels info, remove blank baseline #,7,8,25,45
# #head(RandomVox)
# RandomVox_Zdata <- ddply(RandomVox, .(voxel), transform,  # get z normalized within a voxel across all selected orientations
#                               zSigCh = scale(Hdr))
# #hist(RandomVox_Zdata$zSigCh) # plot histogram
# 
# # writeCSVname
# #savefile <- "s01rFFA_MI_5orient.csv"
```



2nd step: plot hrf of each voxel
```{r plot_hrfperVox, echo=FALSE}
  dev.new()
  ggplot(RandomVox_Zdata, aes(x=volume, y=zSigCh, colour = as.factor(orient))) +
  theme_few() + 
  facet_wrap(~voxel, scales="free_y", nrow=6) + 
  scale_color_brewer(palette='Paired',name="orient",
                       breaks=c("1", "2", "3","4","5","6"),
                       labels=c("L90","L45","Front","R45","R90","180")) + 
  #geom_errorbar(aes(ymin=meanSigCh-se, ymax=meanSigCh+se), width=0.1) +
  geom_point(size=3) +
  geom_line(size=1)
  
```

# kernel density distribution of ROI
```{r BOLD_denisty}

#kdeball <- kde(RandomVox_Zdata$zSigCh)$h # use kde to get bin width

kde_rangeSig <- ddply(RandomVox_Zdata, .(Subj, ROIs), function(dat){
  #print(df$zSigCh)
  #browser()
  rangeSig <- range(dat$zSigCh)
  kk <- kde(dat$zSigCh)$h
  kdeseg <- ceiling(diff(rangeSig)/kk)
  kde_rangeSig <- c(kdeseg,rangeSig)
  
})
colnames(kde_rangeSig) <- c("Subj", "ROIs","kdebins","minRange","maxRange")
RandomVox_Zdata_bins <- merge(RandomVox_Zdata,kde_rangeSig)
# reorder the seq
RandomVox_Zdata_bins <- RandomVox_Zdata_bins[ order(RandomVox_Zdata_bins$Subj, RandomVox_Zdata_bins$ROIs, 
                                                    RandomVox_Zdata_bins$voxel, RandomVox_Zdata_bins$orient, 
                                                    RandomVox_Zdata_bins$volume),]
```

## Definition of entropy
Entropy was a measure introduced by Claude Shannon (1948,"A Mathematical Theory of Communication") to measure the amount of uncertainty in a signal. 

The entropy $H$ is defined as: 
$$
H(X)=-\sum\limits_{x=1}^np(x)\log_2p(x)
$$
$X$ is a discrete random variable and $f(x)$ is the value of its probability distribution at $x$
Because of $log_2$ entropy is measured in bits. 

Entropy in BOLD responses
$$
H(B)=-\sum\limits_{b\in B}p(b)\log_2p(b)
$$

$B$ is the range of BOLD response within a ROI, depends how many bins we defined
Here we caculated the Entroy of BOLD respnse ($H(B)$) of one voxel with 6 orientations (-90, -45, 0, 45, 90, 180)
## need to figure out whether to add 180 (back side of face) ##

3rd step: entropy of the ALL BOLD responses in each voxel
3.1 discretizing the continuous BOLD response into dscrete variable $B$ by dividing the range of responses into a set of equidistant bins $b$  nBins range is based on all voxels within a ROI
3.2 entropy of the BOLD responses in each voxel 4rd step: entropy of the each orientations of a voxel

```{r H(B)_perVox}

computeVox_entropy <- function(Sig, kdebins, minRange, maxRange, MImethod="ML"){
  #browser()
  #BOLD_discreted <- discretize(Sig, numBins=numBins,r=ROI_range)
  BOLD_discreted <- discretize(Sig, numBins=kdebins[1],r=c(minRange[1], maxRange[1]))
  
  HB_perVox <- entropy::entropy(BOLD_discreted, unit="log2", method = MImethod)
  return(HB_perVox)
  }

# H(B)
HB <- ddply(RandomVox_Zdata_bins, .(Subj, ROIs, voxel), summarise, HB_perVox=computeVox_entropy(Sig=zSigCh,kdebins=kdebins,minRange=minRange,maxRange=maxRange, MImethod="ML"))  

#fnEntropy <- function(x) computeVox_entropy(x, kdebins, minRange, maxRange)
#HB <- ddply(RandomVox_Zdata_bins, .(Subj, ROIs, voxel), summarise, HB_perVox=fnEntropy(c(zSigCh,kdebins,minRange, maxRange))) # get entropys of per voxels
# H(B|theta_i)
HBtheta_i <- ddply(RandomVox_Zdata_bins, .(Subj, ROIs, voxel, orient), summarise, HB_perOrientVox=computeVox_entropy(Sig=zSigCh,kdebins=kdebins,minRange=minRange,maxRange=maxRange, MImethod="ML"))   # get entropys of per orient in a voxel

# H(B|theta)

HBtheta <- ddply(HBtheta_i, .(Subj, ROIs, voxel), summarise,
                 meanHBtheta = mean(HB_perOrientVox))


# MI = H(B)-H(B|theta)

MI_table <- merge(HB, HBtheta, by=c("Subj","ROIs","voxel"))

MI_table$MIvalue <- NA
MI_table$MIvalue <- MI_table$HB_perVox - MI_table$meanHBtheta

```


Bias_correct with bootstrapping 
```{r BiasAdj}

shuffled_MI <- function(voxelDat) {
  #browser()
  voxelShuffled <- voxelDat
  voxelShuffled$orient <- sample(voxelShuffled$orient)
  voxelShuffled$volume <- sample(voxelShuffled$volume)
  #HBtheta_i_Shuffled <- ddply(voxelShuffled, .(Subj, ROIs, voxel, orient), 
  #                            summarise, HB_theta_Shuffled=fnEntropy(zSigCh))
  HBtheta_i_Shuffled <- ddply(voxelShuffled, .(Subj, ROIs, voxel, orient), summarise,
                              HB_theta_Shuffled=computeVox_entropy(Sig=zSigCh,kdebins=kdebins,minRange=minRange,maxRange=maxRange, MImethod="ML"))
  
  HB_Shuffled <- ddply(voxelShuffled, .(Subj, ROIs, voxel), summarise, 
                       HB_Shuffled_perVox=computeVox_entropy(Sig=zSigCh,kdebins=kdebins,minRange=minRange,maxRange=maxRange, MImethod="ML"))
  
  #shuf_HB <- computeVox_entropy(voxelShuffled$zSigCh, numBins=numBins, ROI_range = ROI_range)
  meanHBtheta_i_Shuffled <- ddply(HBtheta_i_Shuffled, .(Subj, ROIs, voxel), summarise,
                                  meanHBthetaShuffled=mean(HB_theta_Shuffled))
  
  Shuffled_all <- merge(HB_Shuffled, meanHBtheta_i_Shuffled, by=c("Subj","ROIs","voxel"))
  
  #Shuf_MI <- HB_Shuffled-sum(HBtheta_i_Shuffled$HB_Shuffled_perOrientVox*(1/k)) # We multiply by probability of each orientation (1/k)
  ShuffMI_table <- Shuffled_all$HB_Shuffled_perVox - Shuffled_all$meanHBthetaShuffled
  
  return(ShuffMI_table)
  }

Shuffed_MI_VoxData <- replicate(5000, ddply(RandomVox_Zdata_bins, .(Subj, ROIs, voxel), 
                                         .fun=function(voxelDat) shuffled_MI(voxelDat)), simplify=FALSE)

# list to dataframe
Shuffed_MI_bindVoxData <- do.call(rbind, Shuffed_MI_VoxData) # bind by row
save(Shuffed_MI_bindVoxData, file="s02_MI_5orient_5000permuted.RData")

#hist(Shuffed_MI_bindVoxData$V1)
meanShuffed_MI<- ddply(Shuffed_MI_bindVoxData, .(Subj, ROIs, voxel), summarise,
                       meanShuffed = mean(V1))
#criticalBound <- quantile(meanShuffed_MI$meanShuffed, c(.025, .975))

```

Export MI to CSV file
```{r saveMI2rawdata}

# merge all MI info
RandomVox_Zdata_MI <- merge(RandomVox_Zdata_bins,MI_table)

RandomVox_Zdata_MI_Bias <- merge(RandomVox_Zdata_MI, meanShuffed_MI)

write.csv(RandomVox_Zdata_MI_Bias, file=savefile, row.names=FALSE)

```



```

# shuffled_MIpluginfn <- function(voxelDat) shuffled_MI(voxelDat, k=5, numBins=kdeBins)
# 
# # shufData <- ddply(RandomVox_Zdata, .(Subj, ROIs, voxel), 
# #                   .fun=function(voxelDat) shuffled_MI(voxelDat, k=5, numBins=kdeBins)) # get entropys of per voxels
# 
# Shuffed_MI_VoxData <- replicate(1, ddply(RandomVox_Zdata, .(voxel), 
#                   .fun=function(voxelDat) shuffled_MIpluginfn), simplify=FALSE)
# 
# ddply(RandomVox_Zdata, .(voxel), .fun=function(voxelDat) shuffled_MIpluginfn)


#shufData <- ddply(RandomVox_Zdata, .(voxel), summarise, shuf_MIorient=shuffled_MI(oneVoxelDat, k=5, numBins=kdebins)) # get entropys of per voxels


# shufData <- ddply(RandomVox_Zdata, .(voxel), .fun=function(dat) print(dat)) # get entropys of per voxels


# bootstrapping 1000 times

#kk <- (replicate(1000, shuffled_MI()))
#shufBiasMI <- mean(kk)

#plot(density(kk))

```


Plot hrf of each voxel and sort by MI value
```{r plotMIinROI}

meanMIvalues <- rep(MI_table$MIvalue, each =(nrow(RandomVox_Zdata)/length(unique(RandomVox_Zdata$voxel))))
RandomVox_ZdataMI <- cbind(RandomVox_Zdata,meanMIvalues)
RandomVox_ZdataMI$voxel <- as.factor(RandomVox_ZdataMI$voxel)

RandomVox_ZdataMI$voxel <- reorder(RandomVox_ZdataMI$voxel, -RandomVox_ZdataMI$meanMIvalues) # order voxel by MI values

miLabels <- ddply(RandomVox_ZdataMI, .(voxel), summarise, mi=mean(meanMIvalues)) # get mi label for each voxel
  dev.new()
  ggplot(data=RandomVox_ZdataMI, aes(x=volume, y=zSigCh)) +
  theme_few() + 
  facet_wrap(~voxel, scales="fixed", nrow=6) + 
  scale_color_brewer(palette='Paired',name="orient",
                       breaks=c("1", "2", "3","4","5"),
                       labels=c("L90","L45","Front","R45","R90")) + 
  #geom_errorbar(aes(ymin=meanSigCh-se, ymax=meanSigCh+se), width=0.1) +
  geom_point(aes(colour = as.factor(orient)), size=2) +
  geom_line(aes(colour = as.factor(orient)), size=0.5)  +
  geom_text(data=miLabels, aes(x=20, y=1.5, label=paste(round(mi,2)), group=voxel), size=3)

```


```{r plotHighLowMI}
SortMI<- MI_table[with(MI_table, order(-MIvalue,voxel)),] # sort voxel based on MI value from high to low
HighMIvoxel <- SortMI$voxel[1:1] # get highest MI voxels
LowMIvoxel <- SortMI$voxel[(NROW(SortMI)-1+1):NROW(SortMI)] # get lowest MI voxels
HighMIVox <- droplevels(subset(RandomVox_Zdata, voxel %in% c(HighMIvoxel)))
LowMIVox <- droplevels(subset(RandomVox_Zdata, voxel %in% c(LowMIvoxel)))

  dev.new()
  ghighMI <- ggplot(HighMIVox, aes(x=volume, y=zSigCh, colour = as.factor(orient))) +
  theme_few() + 
  facet_wrap(~voxel, scales="free_y", nrow=6) + 
  scale_color_brewer(palette='Paired',name="orient",
                       breaks=c("1", "2", "3","4","5"),
                       labels=c("L90","L45","Front","R45","R90")) + 
  #geom_errorbar(aes(ymin=meanSigCh-se, ymax=meanSigCh+se), width=0.1) +
  geom_point(size=3) +
  geom_line(size=1) 

  glowMI <- ggplot(LowMIVox, aes(x=volume, y=zSigCh, colour = as.factor(orient))) +
  theme_few() + 
  facet_wrap(~voxel, scales="free_y", nrow=6) + 
  scale_color_brewer(palette='Paired',name="orient",
                       breaks=c("1", "2", "3","4","5"),
                       labels=c("L90","L45","Front","R45","R90")) + 
  #geom_errorbar(aes(ymin=meanSigCh-se, ymax=meanSigCh+se), width=0.1) +
  geom_point(size=3) +
  geom_line(size=1)
  grid.arrange(ghighMI, glowMI, nrow=1)


```


```{r plotDensity}
vox1 <- subset(RandomVox_Zdata, voxel==****) 

vox1$orient <- as.factor(vox1$orient)

ggplot(vox1, aes(x=zSigCh)) + 
    stat_density()


dev.new() 

#Density distribution of a voxel (merge all orients)
g1 <- ggplot(vox1, aes(x=zSigCh)) + 
    theme_few() +
    scale_color_brewer(palette='Set2') + 
    stat_density(aes(group=orient, color=orient, y=0.2*..scaled..), fill=NA, position='identity',bins=kdebins, size=1) + 
    stat_density(fill='grey20', color=NA, size=2, alpha=0.5)

#Density distribution of each orient in a voxel
g2 <- ggplot(vox1, aes(x=zSigCh)) + 
facet_wrap(~orient) + 
 theme_few() + 
 scale_color_brewer(palette='Set2') + 
 stat_density(aes(group=orient, color=orient), fill=NA, position='identity', bins=kdebins, size=1) + 
 stat_density(fill='grey80', color=NA, size=2, alpha=0.5)

#BOLD response of each orient in a voxel
g3 <- ggplot(vox1, aes(x=volume, y=zSigCh))  +
 theme_few() + 
 scale_color_brewer(palette='Set2') + 
 geom_line(aes(color=orient)) + 
 geom_line(stat='summary', fun.y='mean', size=1, alpha=0.5)

grid.arrange(g1, g2, g3, nrow=1)

```



need to do surface, or prediction surface
```{r plot3DScatterMI}
#MIcolvar  <-  color.scale(MI_table$MIvalue,c(0,1,1),c(1,1,0),c(0,1,0))
c=cut(MI_table$MIvalue, breaks=128)
MIcolvar  <-  heat.colors(60)[as.numeric(c)]
open3d()
clear3d("all")
bg3d(color="#887777")
light3d()
plot3d(MI_table$x,MI_table$y, MI_table$z, type='s',col=MIcolvar, size=1, lit=FALSE)

# Fit a plane to the data
fit <- lm( MI_table$z~MI_table$x + MI_table$y)
coefs <- coef(fit)
a <- coefs['MI_table$x']
b <- coefs['MI_table$y']
c <- -1
d <- coefs['(Intercept)']
planes3d(a, b, c, d, alpha=0.2)


#scatter3D(MI_table$x,MI_table$y, MI_table$z, colvar=MI_table$MIvalue )


```



# HBScramble <- NaN
# for (temp in 1:3){
# # H(B)_scramble
#  browser()
#  
#  #fn <- function(x) computeVox_entropy(x)
#  HBScramble[temp] <- ddply(RandomVox_Zdata, .(voxel), summarise, HB_perVox=computeVox_entropy(sample(zSigCh)))
#                              
#  ddply(RandomVox_Zdata, .(voxel), summarise, print(computeVox_entropy(sample(zSigCh))))
#   
#  # H(B|theta_i)_scramble
#  #HBScramble_theta_i(temp) <- ddply(RandomVox_Zdata, .(voxel, orient), summarise, HB_perOrientVox=fn(sample#(zSigCh)))
#   
# }


```

###############
The definition of Mutual information
$$ 
H(X;Y) = -\sum\limits_{x\in X}\sum\limits_{y\in Y}p(x,y)\log_2p(x,y)/p(x)p(y)
H(X;Y) = H(X) - H(X|Y)
= H(Y) - H(Y|X)
= H(X) + H(Y) - H(X,Y)
$$

$H(X), H(Y)$ is marginal entropy of X,Y; $H(X|Y)$ is conditional entropy of X given Y; $H(X,Y)$ is joint entroy of X and Y

Mutual information in BOLD responses
$$
H(B;\Theta) = H(B) - H(B|\Theta)
= H(B) + H(\Theta) - H(B,\Theta)
$$
$\Theta$ is the indvidual orientation that we presented

The definition of conditional entropy
$$
H(Y|X) = -\sum\limits_{x\in X}p(x) H(Y|X = x)
= -\sum\limits_{x\in X}p(x)\sum\limits_{y\in Y}p(y|x)\log_2p(y|x)
= -\sum\limits_{x\in X}\sum\limits_{y\in Y}p(x,y)\log_2p(y|x)
$$

Conditonal Entropy of orientations in BOLD responses
$$
H(B|\Theta) = -\sum\limits_{\theta\in \Theta}p(\theta)\sum\limits_{b\in B}p(b|\theta)\log_2p(b|\theta)
= -\sum\limits_{b\in B}\sum\limits_{\theta\in \Theta}p(\theta,b)\log_2p(b|\theta)
$$
$\Theta$ is the indvidual orientation that we presented, computed condtional entropy $p(b|\theta)$, which measure the resp uncertainty given knowlegde of the stimulus orientation


Since the experimental evaluation of these probabilities is made from a limited number of experimental observa- tions, the resulting information measures suffer from a limited sampling bias (Panzeri et al., 2007). 
Scrambling BOLD response and do bootstrapping

#############################








Reference
http://info-theory.blogspot.jp/2005/01/understanding-mutual-information.html
Gomez-Verdejjo(2012) Analysis of fMRI time series with mutual information 
http://thirdorderscientist.org/homoclinic-orbit/2013/8/25/all-i-really-need-for-data-analysis-i-learned-in-kindergarten-mdash-venn-shannon-and-independence

# to define the bin sizes
Legg(2007)Improving accurary and efficiency of registration by mutual information using Sturges' histogram rule

# to adjust bias
Panzeri S, Senatore R, Montemurro MA, Petersen RS ()

