---
title: "calculate mutual information of each interested region and bootstrapping analysis"
author: "TancyKao"
date: "20150225"
output: html_document
---


```{r loadLibary}
rm(list=ls())
library(plyr)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(ks)
library(entropy)
library(corrplot)
options(device='quartz')
```


```{r loadData}
#setwd("Users/tancy/Dropbox/fMRI_Data_R/fMRI_ERData_R/")
MIdata_s01 <- read.csv('/Users/tancy/Dropbox/fMRIData_R/fMRI_ERData_R/MIData/s01_MI_5orient.csv')
MIdata_s02 <- read.csv('/Users/tancy/Dropbox/fMRIData_R/fMRI_ERData_R/MIData/s02_MI_5orient.csv')
MIdata_s03 <- read.csv('/Users/tancy/Dropbox/fMRIData_R/fMRI_ERData_R/MIData/s03_MI_5orient.csv')
MIdata_s04 <- read.csv('/Users/tancy/Dropbox/fMRIData_R/fMRI_ERData_R/MIData/s04_MI_5orient.csv')
MIdata_s05 <- read.csv('/Users/tancy/Dropbox/fMRIData_R/fMRI_ERData_R/MIData/s05_MI_5orient.csv')

MIdata <- rbind(MIdata_s01,MIdata_s02,MIdata_s03,MIdata_s04,MIdata_s05)
rm(list=c('MIdata_s01','MIdata_s02','MIdata_s03','MIdata_s04','MIdata_s05'))

# rename s02 roi as same as others
MIdata$ROIs <- as.factor(MIdata$ROIs)
MIdata$reROIs <- revalue(MIdata$ROIs, c("arFFA"="rFFA", "prFFA"="rFFA"))

```


The filled bars represent the proportions of neurons in each class that are significantly tuned; that is, for which the mutual information (between neural activity and viewpoint) is higher than the 97.5th percentile of the null distribution. 
Here, we consider shuffed MI values
```{r PercentMI_indROIs}

MIdatavox <- droplevels(ddply(MIdata, .(Subj, reROIs, voxel), summarise,
                              VoxMIvalue = mean(MIvalue),
                              VoxShuffed = mean(meanShuffed)))

MIdataQuant <- droplevels(ddply(MIdatavox, .(Subj, reROIs), summarise,
                                nvox=length(voxel),
                                X975 = quantile(VoxShuffed, 0.85)))

MIdatavoxQuant <- merge(MIdatavox,MIdataQuant)

MIdataSelectvox <- droplevels(ddply(MIdatavoxQuant, .(Subj, reROIs, nvox), summarise,
                                    SelVoxInd=voxel[VoxMIvalue>=X975]))
                                     # get index of those voxel's MIvalue > quantile shuffedMI

freqsMIPerROI <- droplevels(ddply(MIdataSelectvox, .(Subj, reROIs, nvox), summarise, 
                                    freqVox=length(SelVoxInd)))

percentMIPerROI <- droplevels(ddply(freqsMIPerROI, .(Subj, reROIs), transform, 
                                       percentVox = freqVox/nvox*100))



avgSubjMIPerROI <- droplevels(ddply(percentMIPerROI, .(reROIs), summarise,
                                    N=length(Subj),
                                    std=sd(percentVox),
                                    se=std/sqrt(N),
                                    avgpercentVox=mean(percentVox)))
                                     

# plot percentage of orients in a ROI
dev.new()
ggplot(percentMIPerROI, aes(x=reROIs, y=percentVox)) + 
  facet_wrap(~Subj, scales='free') + 
  geom_bar(stat="identity",colour="black")


# plot percentage of orients in a ROI
dev.new()
ggplot(avgSubjMIPerROI, aes(x=reROIs, y=avgpercentVox)) + 
  geom_errorbar(aes(ymin=avgpercentVox-se, ymax=avgpercentVox+se), width=0.1) +
  geom_bar(stat="identity",colour="black")


```


Presenting BOLD response and MI and compare the ROIs
```{r maxSigChVsMI}
meanSigChOrientTruncate <- droplevels(subset(MIdata, (volume >= 2) & (volume <=10)))
MIdataSigChvox <- droplevels(ddply(MIdata, .(Subj, reROIs, voxel, orient), summarise,
                                   maxSig=max(zSigCh),
                                   VoxMIvalue = mean(MIvalue),
                                   VoxShuffed = mean(meanShuffed)))

MIdataSigChvoxWcut <- merge(MIdataSigChvox, MIdataQuant)

maxSigPerVoxCut <- ddply(MIdataSigChvoxWcut, .(Subj, reROIs, voxel), summarise,
                      avgSig=mean(maxSig),
                      VoxMIvalue=mean(VoxMIvalue),
                      VoxShuffed=mean(VoxShuffed),
                      cutX975=mean(X975))
#maxSigPerVoxCut$cutX975 <- 0
maxSigPerVox <- ddply(maxSigPerVoxCut, .(Subj, reROIs, voxel), summarise,
                      SelVoxInd=voxel[VoxMIvalue>=cutX975],
                      avgSig=avgSig[VoxMIvalue>=cutX975],
                      avgMI=VoxMIvalue[VoxMIvalue>=cutX975])

maxSigPerVox$mergLHRH <- revalue(maxSigPerVox$reROIs, c("rFFA"="FFA", "lFFA"="FFA","rOFA"="OFA","lOFA"="OFA"))                      

maxSigPerVox.LH <- subset(maxSigPerVox, reROIs==c('lFFA','lOFA'))
maxSigPerVox.RH <- subset(maxSigPerVox, reROIs==c('rFFA','rOFA'))

maxSigPerVox.LH$reROIs <- factor(maxSigPerVox.LH$reROIs) # drop levels after subset
maxSigPerVox.RH$reROIs <- factor(maxSigPerVox.RH$reROIs) # drop levels after subset

# plot percentage of orients in a ROI
# merge left and right ROIs
dev.new()
ggplot(maxSigPerVox, aes(x=avgMI, y=avgSig, colour=mergLHRH)) + 
  facet_wrap(~Subj, scales='free') + 
  geom_point()
  #stat_summary(aes(group = mergLHRH), fun.y=mean, geom="point") 

# all ROIs
dev.new()
ggplot(maxSigPerVox, aes(x=avgMI, y=avgSig, colour=reROIs)) + 
  facet_wrap(~Subj, scales='free') + 
  geom_point()

# only Left 
dev.new()
ggplot(maxSigPerVox.LH, aes(x=avgMI, y=avgSig, colour=reROIs)) + 
  facet_wrap(~Subj, scales='free') + 
  geom_point()

# only Right
dev.new()
ggplot(maxSigPerVox.RH, aes(x=avgMI, y=avgSig, colour=reROIs)) + 
  facet_wrap(~Subj, scales='free') + 
  geom_point()


maxSigPerROI <- ddply(maxSigPerVox, .(Subj, reROIs), summarise,
                      N=length(voxel),
                      std=sd(avgSig),
                      se=std/sqrt(N),
                      avgMI=mean(avgSig))

 


maxSigPerROI.RH <- subset(maxSigPerROI, reROIs==c('rFFA','rOFA'))
maxSigPerROI.LH <- 

#
dev.new()
ggplot(maxSigPerROI.RH, aes(x=reROIs, y=avgMI, fill=reROIs)) + 
  facet_wrap(~Subj, scales='free') + 
  geom_bar(fill="#DD8888", width=.8, stat="identity") +
  geom_errorbar(aes(ymin=avgMI-se, ymax=avgMI+se), width=0.1) +
  guides(fill=FALSE) +
  xlab("face-selective areas") + ylab("Neural Responses") 




```

```{r plotDesityMIandBOLD}
# plot density of MI and BOLD response of two rois

OneSubjPlot <- droplevels(subset(maxSigPerVox.RH, Subj=="S05"))

# placeholder plot - prints nothing at all
empty <- ggplot() + geom_point(aes(1, 1), colour = "white") + theme(plot.background = element_blank(), 
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
    panel.border = element_blank(), panel.background = element_blank(), axis.title.x = element_blank(), 
    axis.title.y = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank(), 
    axis.ticks = element_blank())

# scatterplot of x and y variables
scatter <- ggplot(OneSubjPlot, aes(x=avgMI, y=avgSig)) + 
  geom_point(aes(color = reROIs)) + 
  scale_color_manual(values = c("orange", "purple")) + 
  theme(legend.position = c(1, 1), legend.justification = c(1,1))

# marginal density of x - plot on top
plot_top <- ggplot(OneSubjPlot, aes(x=avgMI, fill = reROIs)) + 
  geom_density(alpha = 0.5) + 
  scale_fill_manual(values = c("orange", "purple")) + 
  theme(legend.position = "none")

# marginal density of y - plot on right
plot_right <- ggplot(OneSubjPlot, aes(x=avgSig, fill = reROIs)) + 
  geom_density(alpha = 0.5) + 
  coord_flip() + 
  scale_fill_manual(values = c("orange", "purple")) + 
  theme(legend.position = "none")

# arrange the plots together, with appropriate height and width for each row
# and column
grid.arrange(plot_top, empty, scatter, plot_right, ncol = 2, nrow = 2, widths = c(4,1), heights = c(1, 4))


```


Bootstrap T-test between two MI values of reROIs
```{r Bootstrap}
library(boot)

diff_boot = function(d1,i){
  #browser()
  d = d1
  d$reROIs <- d$reROIs[i] # randomly re-assign groups
  Mean = tapply(d$avgMI, d$reROIs, mean)
  Diff = Mean[1]-Mean[2]
  Diff
}


diff_sigCh = function(d1,i){
  #browser()
  d = d1
  d$reROIs <- d$reROIs[i] # randomly re-assign groups
  Mean = tapply(d$avgSig, d$reROIs, mean)
  Diff = Mean[1]-Mean[2]
  Diff
}



# individual Subj bootsrap test
IDs=unique(maxSigPerVox.RH$Subj)
for (j in seq(1, length(IDs))){
  
  OneSubjMI <- subset(maxSigPerVox.RH, Subj==IDs[j])
  OneSubjMI$Subj <- factor(OneSubjMI$Subj) # drop levels after subset
  Boot_result = boot(data=OneSubjMI, statistic = diff_boot, R =1000)
  p_value<- mean(abs(Boot_result$t) > abs(Boot_result$t0)) # p value
  print(c(j, p_value))
  #quantile(Boot_result$t, c(.025, .975))
  #Boot_result$t0
}


# Across all subjects
Boot_result = boot(data = maxSigPerVox.LH, statistic = diff_sigCh, R =10000)
  p_value_acrSubj <- mean(abs(Boot_result$t) > abs(Boot_result$t0)) # p value
  p_value_acrSubj
  #hist(Boot_result$t)
  #plot(density(Boot_result$t))
  #quantile(Boot_result$t, c(.025, .975))


# for merge left and right sides
diff_mergeboot = function(d1,i){
  #browser()
  d = d1
  d$mergLHRH <- d$mergLHRH[i] # randomly re-assign groups
  Mean = tapply(d$avgMI, d$mergLHRH, mean)
  Diff = Mean[1]-Mean[2]
  Diff
}

mergeB_result = boot(data = maxSigPerVox, statistic = diff_mergeboot, R =10000)
  p_value_merge <- mean(abs(mergeB_result$t) > abs(mergeB_result$t0)) # p value
  p_value_merge




```


***** NO RUN *****



f <- function() {
   A <- sample(maxSigPerVox$avgMI, nrow(maxSigPerVox$reROIs), replace=T) 
   B <- sample(maxSigPerVox$avgMI, nrow(maxSigPerVox), replace=T)
   mean(A)-mean(B)
}

kk <- (replicate(1000, f()))
hist(kk)
quantile(kk, c(.025, .975))
plot(density(kk))





Kullback-Leibler divergence (information divergence)
non-symmetric measure of the difference between two probability distributions P and Q
$$
DKL(\P||\Q) = \Sigma{i}\P(\i)\log_2/P(\i)/Q(\i)
$$



Calculate KL (not consider shuffed MI)
```{r KL_divergence}

orKL_fun <- function(voxelDat){ 
  #browser()
  tempVox <- voxelDat
  or1 <- droplevels(subset(tempVox, orient==1))
  or2 <- droplevels(subset(tempVox, orient==2))
  or3 <- droplevels(subset(tempVox, orient==3))
  or4 <- droplevels(subset(tempVox, orient==4))
  or5 <- droplevels(subset(tempVox, orient==5))
  
  kdeOr1 <- kde(or1$zSigCh)
  kdeOr2 <- kde(or2$zSigCh)
  kdeOr3 <- kde(or3$zSigCh)
  kdeOr4 <- kde(or4$zSigCh)
  kdeOr5 <- kde(or5$zSigCh)
  
  or11_KL <- KL.empirical(kdeOr1$estimate, kdeOr1$estimate, unit="log2")
  or12_KL <- KL.empirical(kdeOr1$estimate, kdeOr2$estimate, unit="log2")
  or13_KL <- KL.empirical(kdeOr1$estimate, kdeOr3$estimate, unit="log2")
  or14_KL <- KL.empirical(kdeOr1$estimate, kdeOr4$estimate, unit="log2")
  or15_KL <- KL.empirical(kdeOr1$estimate, kdeOr5$estimate, unit="log2")
  or1_allKL <- c(or11_KL,or12_KL,or13_KL,or14_KL,or15_KL)
  
  or22_KL <- KL.empirical(kdeOr2$estimate, kdeOr2$estimate, unit="log2")
  or23_KL <- KL.empirical(kdeOr2$estimate, kdeOr3$estimate, unit="log2")
  or24_KL <- KL.empirical(kdeOr2$estimate, kdeOr4$estimate, unit="log2")
  or25_KL <- KL.empirical(kdeOr2$estimate, kdeOr5$estimate, unit="log2")
  or2_allKL <- c(or12_KL,or22_KL,or23_KL,or24_KL,or25_KL)
  
  or33_KL <- KL.empirical(kdeOr3$estimate, kdeOr3$estimate, unit="log2")
  or34_KL <- KL.empirical(kdeOr3$estimate, kdeOr4$estimate, unit="log2")
  or35_KL <- KL.empirical(kdeOr3$estimate, kdeOr5$estimate, unit="log2")
  or3_allKL <- c(or13_KL,or23_KL, or33_KL,or34_KL,or35_KL)
  
  or44_KL <- KL.empirical(kdeOr4$estimate, kdeOr4$estimate, unit="log2")
  or45_KL <- KL.empirical(kdeOr4$estimate, kdeOr5$estimate, unit="log2")
  or4_allKL <- c(or14_KL,or24_KL,or34_KL, or44_KL,or45_KL)
  
  or55_KL <- KL.empirical(kdeOr5$estimate, kdeOr5$estimate, unit="log2")
  or5_allKL <- c(or15_KL,or25_KL,or35_KL, or45_KL, or55_KL)
  
  OrKL_all <- rbind(or1_allKL,or2_allKL,or3_allKL,or4_allKL,or5_allKL)
  colnames(OrKL_all) <- c("orL90", "orL45","Front","orR45","orR90")
  rownames(OrKL_all) <- c("orL90", "orL45","Front","orR45","orR90")
  return(OrKL_all)  
  }

# find the best MI voxels in each roi
cutMIcrit=c(0, 0.25, 0.5, 0.75)
MIpercentil <- droplevels(ddply(MIdata, .(Subj, reROIs), summarise,
                                Xcut = quantile(MIvalue, cutMIcrit[4])))

MIdata.MIpercentil <- merge(MIdata, MIpercentil)
MIdataBestMIVox <- droplevels(subset(MIdata.MIpercentil, MIvalue>=Xcut)) # select those voxels > critical

MI_KLdata <- ddply(MIdataBestMIVox, .(Subj, reROIs, voxel), 
                   .fun=function(voxelDat) orKL_fun(voxelDat))

# add one column as orientation index
MI_KLdata$orIndex <- rep(seq(1,5), nrow(MI_KLdata)/5)

# avg all voxels in a roi
avgKLdata <- ddply(MI_KLdata, .(Subj, reROIs, orIndex), summarise,
                      mean_orL90 = mean(orL90),
                      mean_orL45 = mean(orL45),
                      mean_orFront = mean(Front),
                      mean_orR45 = mean(orR45),
                      mean_orR90 = mean(orR90))


# plot KL of avg voxels in reROIs
j=0
for (i in seq(1,nrow(avgKLdata),5)){
  j=j+1
  ROIname <- unique(avgKLdata$reROIs)
  avgKL.matrix <- as.matrix(avgKLdata[c(4:8)])
  avgKLROI.matrix <- avgKL.matrix[i:(i+4),1:5]
  dev.new() 
  orTitle=ROIname[j]
  corrplot(avgKLROI.matrix,title=orTitle, cl.lim=c(0,max(avgKLdata[c(4:8)])+0.05), method="circle", 
           is.corr=FALSE, type="full", addCoef.col="black",
           addshade="positive", addgrid.col="gray", mar=c(0,0,1,0), diag= FALSE) 
  }

```






corrplot(tempMergeOrKL, method="shade", shade.col=NA, tl.col="black",tl.srt=45, addshade="positive")


dev.new() 
orTitle="lFFA"
corrplot(OrKL_all,title=orTitle, cl.lim=c(0,1), method="color", is.corr=FALSE, type="full",
addshade="positive", addgrid.col="gray", mar=c(0,0,1,0), diag= FALSE) 

