---
title: "erfMRI_plotData"
author: "TancyKao"
date: "20150510"
output: html_document
---

Initializing and loading libraries
```{r Init}
## read matlab data to R
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(plyr)
library(reshape)

options(device='quartz')
# recall data

#setwd("~/Dropbox/02_RikenProj/fMRI_Data_R/")
allData <- read.csv('/Users/tancy/Dropbox/fMRIData_R/fMRI_ERData_R/SigCh_OrientData.csv')

```

############################ ROIs plot ####################################
Plot r2 distribution by ROIs
```{r r2_Distribution,  fig.width=10, fig.height=7}
r2Voxel <- droplevels(ddply(allData, .(Subj, ROIs, voxel), summarise,
                            mean_r2 = mean(r2)))

dev.new()
ggplot(r2Voxel, aes(x = mean_r2)) + 
  facet_wrap(Subj ~ROIs, scales='free') + 
  geom_histogram(binwidth = 0.02, fill="white",colour="black")

```


Plot hrf of Orient by ROIs
```{r plot_hrf, fig.width=10, fig.height=7}
## get voxels > r2
allDataSub <- subset(allData, r2 > 0.06 & ROIs!='rOTA' & ROIs!='lOTA')

## get mean signal of each orientation
#  only treat ROIs, orient, volume, Side as variables
meanSigPerOrient <- droplevels(ddply(allDataSub, .(Subj, ROIs, orient, volume), summarise, 
                                     N = length(voxel),
                                     std=sd(SignalChange),
                                     se=std/sqrt(N), 
                                     meanSigCh = mean(SignalChange))) # avoid to use same name

#allDataSub$orient <- as.factor(allDataSub$orient)
## plot hrf curve of mean signal change (volume by mean signal change), ROIs as category (meanSigPerOrient)
ggplot(meanSigPerOrient, aes(x=volume, y=meanSigCh, colour = as.factor(orient))) +
  theme_few() + 
  facet_wrap(Subj ~ROIs) +
  scale_color_brewer(palette='Paired',name="orient",
                       breaks=c("1", "2", "3","4","5","6","7"),
                       labels=c("L90","L45","Front","R45","R90","180","Baseline")) + 
  geom_errorbar(aes(ymin=meanSigCh-se, ymax=meanSigCh+se), width=0.1) +
  geom_point(size=1) +
  geom_line(size=0.5)

```

Plot maxSignalChange of Orient by ROIs; 
Each points denote voxels
```{r plot_maxSigCh}
# max value of each orient

meanSigPerOrientTruncate <- droplevels(subset(meanSigPerOrient, (volume >= 2) & (volume <=15)))
maxSigPerVox <- ddply(meanSigPerOrientTruncate, .(Subj, ROIs, orient), summarise,
                      maxSig=max(meanSigCh),
                      se =se[meanSigCh==maxSig],
                      volume = volume[meanSigCh==maxSig])
unique(maxSigPerVox$volume)

maxSigPerVox$ROIs <- as.factor(maxSigPerVox$ROIs)
maxSigPerVox$reROIs <- revalue(maxSigPerVox$ROIs, c("arFFA"="rFFA", "prFFA"="rOFA"))
                             
#maxSigPerVox$ROIs[maxSigPerVox$ROIs %in% c("arFFA")] <- "rFFA"
#maxSigPerVox$ROIs[maxSigPerVox$ROIs %in% c("prFFA")] <- "prFFA"

avgSubjmaxSig <- ddply(maxSigPerVox, .(reROIs, orient), summarise,
                       N=length(Subj),
                       std=sd(maxSig),
                       se=std/sqrt(N),
                       avgMaxSigCh = mean(maxSig))

## multiple histograms of max signal change (orient by maxSig), ROIs as category
dev.new()
ggplot(maxSigPerVox, aes(x=orient, y=maxSig)) + 
  theme_few() + 
  facet_wrap(Subj ~ROIs, scales='free') + 
  geom_smooth(aes(group=1), method='loess', se=FALSE) + 
  geom_errorbar(aes(ymin=maxSig-se, ymax=maxSig+se), width=0.1) +
  #geom_point(size=5)
  geom_bar(stat='identity')

# plot avg of all subjects
dev.new()
ggplot(avgSubjmaxSig, aes(x=orient, y=avgMaxSigCh)) + 
  theme_few() + 
  facet_wrap(~reROIs, scales='free') + 
  geom_smooth(aes(group=1), method='loess', se=FALSE) + 
  geom_errorbar(aes(ymin=avgMaxSigCh-se, ymax=avgMaxSigCh+se), width=0.1) +
  #geom_point(size=5)
  geom_bar(stat='identity')


```

plot stack bars (percentage of each orient in ROI) 
```{r plot_RatioMaxSigChOrient}
allDataSubTruncate <- droplevels(subset(allDataSub, (volume >= 1) & (volume <=15) & orient!=7 & orient!=6))

maxSigPerOrientPerVox <- droplevels(ddply(allDataSubTruncate, .(Subj, ROIs, voxel), summarise, 
                                       maxSigPerOrient=max(SignalChange),
                                       orient=orient[SignalChange==maxSigPerOrient]))

maxSigPerOrientPerVox$ROIs <- as.factor(maxSigPerOrientPerVox$ROIs)
maxSigPerOrientPerVox$reROIs <- revalue(maxSigPerOrientPerVox$ROIs, c("arFFA"="rFFA", "prFFA"="rOFA"))


freqSigOrientPerVox <- droplevels(ddply(maxSigPerOrientPerVox, .(reROIs, orient), summarise, 
                                       freqOrient=length(orient)))

percentSigOrient <- droplevels(ddply(freqSigOrientPerVox, .(reROIs), transform, 
                                       percentOrient = freqOrient/sum(freqOrient)*100))

# get y position of label of percentage
percentSigOrient <- droplevels(ddply(percentSigOrient, .(reROIs), transform, 
                                       label_y = cumsum(percentOrient)-0.5*percentOrient)) 


# plot percentage of orients in a ROI
dev.new()
ggplot(percentSigOrient, aes(x=reROIs, y=percentOrient, fill= as.factor(orient))) + 
  geom_bar(stat="identity",colour="black") + 
  guides(fill=guide_legend(reverse=TRUE)) +
  #scale_fill_gradient(low='white', high='grey20')
  scale_fill_brewer(palette="Pastel1") +
  geom_text(aes(y=label_y, label=round(percentOrient)),colour="black")

```


############################ voxel plot #############################################

```{r plot_maxSigChVoxel}
## get z score
#List_SubjROIs <- split(allDataSub, allDataSub$Subj %in% allDataSub$ROIs) # split the dataset into a list of datasets based on the subjects & ROIs
    
#SubROIdata <- droplevels(subset(allDataSub, (Subj=="S02") & (ROIs %in% c('arFFA', 'prFFA', 'rOFA', 'rOTA'))))


# transferm signal change of each voxel to z-score and add to last column
SubROIZdata <- ddply(allData, .(Subj, voxel, ROIs), transform, 
                     zSigCh = scale(SignalChange))

SubROIdataTruncate <- droplevels(subset(SubROIZdata, (volume >= 1) & (volume <=15)))
#maxSigdata <- ddply(SubROIdataTruncate, .(Subj, voxel, orient, ROIs), summarise,  
#                 maxSigCh = max(SignalChange))
    
maxZSigdata <- ddply(SubROIdataTruncate, .(Subj, voxel, orient, ROIs), summarise,  
                     maxSigCh = max(SignalChange),
                     maxZSigCh = max(zSigCh))



## plot peak resp of orient in each voxel
for (ixSubj in unique(maxZSigdata$Subj) ) {
  #print(ixSubj)
  dev.new()
  #browser()
  oneSubjData <- droplevels(subset(maxZSigdata, Subj==ixSubj))
  gg <- ggplot(oneSubjData, aes(x=orient, y=maxZSigCh)) + 
    facet_grid(~ROIs, scales='free') +
    ggtitle(paste('Subject:', ixSubj)) +
    theme_few() +
    geom_point() +
    geom_smooth(aes(group=1),method='loess', se=FALSE) 
  print(gg)
}
```

## plot heat map for hrf of each voxel
```{r plotHeatMap}
IndData <- droplevels(subset(SubROIZdata, (Subj=="S02") & ROIs=="rOFA" & voxel %in% c(1,2)))

myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")), space="Lab")
ggplot(IndData, aes(x=volume, y=orient, fill=(zSigCh))) + 
  theme_few() +
  facet_grid(~voxel, scales='free',space = "free" ) + 
  geom_tile(aes(fill=(zSigCh))) + 
  scale_fill_gradientn(colours = myPalette(100)) 

```

## correlation 
maxZdataWide <- cast(maxZSigdata, orient~ROIs+voxel, value='maxZSigCh')
cors <- melt(cor(maxZdataWide))
tt <- subset(cors, value<1)
tt <- tt[order(-tt$value),]
roiOrder <- with(tt, reorder(X1, value))
tt$X1 <- factor(tt$X1, levels=levels(roiOrder))

tt1 <- tt[order(-tt$value),]


ggplot(data=tt, aes(x=X1, y=X2, fill=value)) + 
  theme_few() + 
  theme(axis.text.x=element_text(angle=45, hjust=1)) + 
  coord_equal() + 
  scale_fill_gradientn(colours = myPalette(100)) + 
  geom_tile()

###########################################################################
## Orientation tuning index (OSI, based on Swindale 1998 paper)
## Tuning function of each neuron. The tuning function is defined using 
## vector sum

```{r OSI_for_each_roi}
SubROImaxdata <- droplevels(subset(maxZSigdata, Subj %in% c('S03') & 
                                     orient %in% c("1","2","3","4","5") & ROIs %in% c('rFFA')))
SubROImaxdata <- droplevels(subset(maxZSigdata, orient %in% c("1","2","3","4","5")))

maxZsigdata_rmOTA <- droplevels(subset(maxZSigdata, ROIs!= "rOTA" & ROIs!= "lOTA"))
 
maxZsigdata_rmOTA <- droplevels(subset(maxZSigdata, ROIs %in% c("rFFA")))
 

for (ixSubj in unique(maxZsigdata_rmOTA$Subj) ) 
  {
    for (ixROIs in unique(maxZsigdata_rmOTA$ROIs) )
      {
        SubROImaxdata <- droplevels(subset(maxZsigdata_rmOTA, Subj==ixSubj & ROIs==ixROIs & 
                                             orient %in% c("1","2","3","4","5")))  

        sdat  <- with(SubROImaxdata, split(SubROImaxdata, voxel)) # split data by voxel

        ## OSI, based on Swindale 1998 paper
        OSI_perVox <- function(RespOrient,presentOrient)
          {
                OSI_equation <- sqrt((sum(RespOrient*cos(2*presentOrient))^2)+
                                    (sum(RespOrient*sin(2*presentOrient))^2))/sum(RespOrient)
                #browser()
                return(OSI_equation)  
          }

        presentOrient <- c(-90,-45,0,45,90)
        #browser()
        print(ixSubj)
        print(ixROIs)
        
        RespOrient <- sapply(1:length(sdat),function(x) t(sdat[[x]]$maxZSigCh)) # t function is to transpose
        RespOrient <- data.frame(RespOrient)
        
        #View(RespOrient)
        #colnames(RespOrient) <- presentOrient
        OSIdata <- sapply(RespOrient, OSI_perVox, presentOrient=presentOrient)
        OSIdata <- data.frame(OSIdata)
        
        dev.new()
        gg <- ggplot(OSIdata, aes(x=OSIdata)) + 
              #facet_wrap(~ROIs, scales='free') + 
              ggtitle(paste('Subject:', ixSubj, 'ROIs:', ixROIs)) +
              theme_few() +
              #xlim(c(0, 1)) +
              #ylim(c(0, 1)) +
              geom_histogram(binwidth = 0.05, fill="white",colour="black")
             print(gg)
         }
    }

#prcRunsPerModel$SubjectID <- factor(prcRunsPerModel$SubjectID, levels=levels(subjectOrder))
#subjectOrder <- with(nRunsPerModelWide, reorder(SubjectID, -prcFullModelRuns))


```
